2.4 PRNT50

### 2025-08-12 2_scripts/2.3 Prob Model for PCR pos.qmd
Minor updates

```{r}
# How gaps were calculated: 
# X: Oct 2023, btw raw CBS & APL, Y: APL vs transformed CBS, Z: CBS vs transformed APL
tb_alberta %>%
  select(data, samplemonth, prop, raw) %>%
  .[.$samplemonth == as.yearmon("Oct 2022"),] %>%
  mutate(X = .$prop[4] - .$prop[1]) %>%
  mutate(Y = .$prop[3] - .$prop[1]) %>%
  mutate(Z = .$prop[4] - .$prop[2])
```

@@

#### 3.6 Create fig panel for the manuscript
```{r}
# use cowplot package
p08.upper<-cowplot::plot_grid(p03.5  + ylab("Seroprevalence") +
                              ylim(0,0.8), 
                              p03.3  + ylab("") +
                              ylim(0,0.8)+
                              scale_y_continuous(breaks = c(0,0.2,0.4,0.6,0.8)),
                              nrow=1, align = "h",
                              labels = c("A", "B"))
p08.upper

@@

##### 4.3.1 Create monthly anti-N positivity data for CBS
```{r}
# CBSs AB subset, N=124008
CBSs_AB<-CBSs[CBSs$province == "AB",]

@@

##### 4.3.2 Create monthly anti-N positivity data for APL
```{r}
# APL data, N=214776

@@

##### 4.3.3 Make plots
```{r}
p4.3.3<-tb_43m %>%
  ggplot(aes(x=samplemonth, y=prop, colour=study)) +
  geom_point(aes(colour=study)) +
  geom_line(aes(colour=study, linetype=adjustment))  +
  xlab("Month of Sample Collection") +
  ylab("Seroprevalence") +
  coord_cartesian(xlim = c(as.yearmon("May 2020"), as.yearmon("Jan 2022")),
                  ylim = c(0, 0.15)) +
#  ylim(c(as.yearmon("Jan 2022"), as.yearmon("Feb 2023"))) +
  scale_colour_manual("", values = c("red","blue")) +
  scale_linetype_manual("", values=c(1,2,3),
                        labels= c("Unadjusted", "RG adjusted (informed priors)",
                                  "RG adjusted (nautral priors)")) +
  theme_bw()

>>>>

```{r}
# What is the numbers in 2022 Oct?
tb_alberta %>%
  filter(samplemonth == as.yearmon("2022-10-01"))

# How gaps were calculated: 
# X: Oct 2023, btw raw CBS & APL, Y: APL vs transformed CBS, Z: CBS vs transformed APL
tb_alberta %>%
  select(data, samplemonth, prop, raw) %>%
  filter(samplemonth == as.yearmon("2022-10-01")) %>%
  summarise(
    A = prop[1], B = prop[4],
    X = .$prop[4] - .$prop[1],
    Y = .$prop[3] - .$prop[1],
    Z = .$prop[4] - .$prop[2]
  )

@@

#### 3.6 Create fig panel for the manuscript

```{r}
# use cowplot package
p08.upper<-cowplot::plot_grid(p03.5  + ylab("Seroprevalence") +
                              ylim(0,0.8), 
                              p03.3  + ylab("") +
                              ylim(0,0.8)+
                              scale_y_continuous(breaks = c(0,0.2,0.4,0.6,0.8)),
                              nrow=1, align = "h",
                              labels = c("A", "B"))
p08.upper

@@

##### 4.3.1 Create monthly anti-N positivity data for CBS

```{r}
# CBSs AB subset, N=124008
CBSs_AB<-CBSs[CBSs$province == "AB",]

@@

##### 4.3.2 Create monthly anti-N positivity data for APL

```{r}
# APL data, N=214776

@@

##### 4.3.3 Make plots

```{r}
p4.3.3<-tb_43m %>%
  ggplot(aes(x=samplemonth, y=prop, colour=study)) +
  geom_point(aes(colour=study)) +
  geom_line(aes(colour=study, linetype=adjustment))  +
  xlab("Month of Sample Collection") +
  ylab("Seroprevalence") +
  coord_cartesian(xlim = c(as.yearmon("May 2020"), as.yearmon("Jan 2022")),
                  ylim = c(0, 0.15)) +
#  ylim(c(as.yearmon("Jan 2022"), as.yearmon("Feb 2023"))) +
  scale_colour_manual("", values = c("red","blue")) +
  scale_linetype_manual("", values=c(1,2,3),
                        labels= c("Unadjusted", "RG adjusted (informed priors)",
                                  "RG adjusted (nautral priors)")) +
  theme_bw()

### 2025-08-12 2_scripts/2.4 PRNT50.qmd
All figs match w the ver in the rev_1
Minor updates

#### 2.2 Roche-S --> Abbott-S transformation
This part is skipped and not mentioned in the manuscript.

@@

```{r}
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, group=data, colour=data)) +
  geom_point(aes(colour=data)) +
  geom_line(aes(colour=data))  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red", "purple"),
                      labels = c("AB General Pop.", "APL Vax%", 
                                 "Anti-S(+)", "Anti-S(+) N(-)")
                      )
#  scale_linetype_manual("", values=c(3,3,1,1), labels = p09.lab) 
p12
```
#### 4.2.4 Apply (S+ N-)/N- for vax% estimation

First we calculate (S+ N-)/N- among CBS samples
```{r}
table(CBSs_AB$pos_roche, CBSs_AB$pos_roche_s, useNA = 'ifany')
# good data quality, only two NAs in roche_n and roche_s

@@

Then we calculate the vax% using S(+) and (S+ & N-)/(N-)
```{r}
tbCBS_vax_est<-CBSs_AB %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            n_pos = sum(pos_roche, na.rm = T),
            s_pos = sum(pos_roche_s, na.rm = T),    
            n_vax= sum(vax_est, na.rm = T)
            )
tbCBS_vax_est$n_neg = tbCBS_vax_est$n - tbCBS_vax_est$n_pos
tbCBS_vax_est$prop.vax<-tbCBS_vax_est$n_vax/tbCBS_vax_est$n_neg

@@

#### Now redo the plot
```{r}
# Add study and adjustment
dta.vax$study[dta.vax$data %in% c("AB_general")]<-"AB_general"
dta.vax$study[dta.vax$data %in% c("CBS: S_pos_N_Neg")]<-"CBS"
dta.vax$study[dta.vax$data %in% c("APL_vax", "S_pos", "S_pos_N_Neg")]<-"APL"

@@

p12.lab <- c("AB General Pop.", "APL Vax%", 
             "APL: Anti-S(+)", "APL: Anti-S(+) N(-)", "CBS: Anti-S(+) N(-)")
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, colour=interaction(study, adj),
                                        linetype=interaction(study, adj))) +
  geom_point() +
  geom_line()  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red","purple", "blue"),
                      labels = p12.lab)+
  scale_linetype_manual("", values=c(1,1,3,3,3), labels = p12.lab) +
  theme(legend.position = c(0.8,0.25))
p12
```
#### Print p12
```{r}
png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1440, res=300)
p12
dev.off()
```
#### 4.2.5 Figure S2 Panel B using transformed data
```{r}
# APL S+/N- N is transformed (Abbott-->Roche)

@@

##### Print p12 and p13 together
```{r}
p1213<-plot_grid(p12 + xlab ("Time of Sample Collection") +
                 coord_cartesian(xlim=as.yearmon(c("Oct 2020", "Feb 2023")))+
                 theme(legend.position = c(0.67,0.25)), 
                 p13 + xlab ("Time of Sample Collection") + ylab("") +
                 coord_cartesian(xlim=as.yearmon(c("Oct 2020", "Feb 2023")))+
                 theme(legend.position = c(0.65,0.2)),
                 labels = c("A", "B"),
                 nrow=1,  align = "h")
p1213

png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1200, res=300)
p1213
dev.off()
```
##### Print p12 and p13 in pdf formats, as requested by AJE
```{r}
ggsave(file="../4_output/figs/Fig.S2 Vax.pdf", 
       width=2400, height=1200, dpi=300, units = "px", 
       plot = p1213)
```

@@

#### 4.4.1 add prop (1-dose only)
```{r}
# number of personals in each month
tb_abbsvall<-APLs %>%
  .[!(.$vax %in% c("No_ULI")),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsvall$n.all<-tb_abbsvall$n

@@

#### 4.4.2 Making the plot:
```{r}
p09v1a<-tb_abbsv1 %>%
  ggplot(aes(x=samplemonth)) +
  geom_point(aes(y=prop.abbs1), colour="red") +
  geom_point(aes(y=prop.abbs2), colour="red") +
  geom_line(aes(y=prop.abbs1), colour="red")  +
  geom_line(aes(y=prop.abbs2), linetype=2, colour="red")  +
  geom_line(aes(y=prop.d1), colour="darkgrey") +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: 1 dose only")
p09v1a
```

#### 4.4.3 Check the time gap btw 1st dose and 2nd dose. 
```{r}
APL_vax<-RFD4682_v2_vaxe %>%
  select(ULI, ULI_IDe, DOSE_COUNT, IMMUNIZATION_DATE) %>%
  .[.$DOSE_COUNT==1 | .$DOSE_COUNT==2, ]

@@

summary(as.numeric(APL_vax$time_gap))
```
##  4.5 Adding the proportion of each stratum

@@

#### The unvaccinated stratum
```{r}
# add proportion of the stratum into the data table
tb_abbsv0<-merge(tb_abbsv0, select(tb_abbsv1, samplemonth, n.all),
      by="samplemonth", all.x=T, all.y=F)
tb_abbsv0<-tb_abbsv0 %>%
  mutate(
    prop.stra = n/n.all
  )
p09v0.lab<-c("APL (manufacturer cutoff)", "APL (PRNT-50 cutoff)",
             "Proportion of the stratum")

@@

#### The 1-dose only stratum
```{r}
tb_abbsv1_melt<-reshape2::melt(select(tb_abbsv1, -c(2:4,7)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")

@@

#### The 2-dose+ stratum
```{r}
tb_abbsv2<-tb_abbsv2[-1,]

@@

### merge all four into one figue panel
```{r}
rm(p09v0a, p09v1a, p09v2a)
```

#### Figure panel
```{r}
# use cowplot package
p09.upper<-cowplot::plot_grid(p09c + ylab("Proportion") + xlab ("") +
                           coord_cartesian(xlim=as.yearmon(c("Jan 2021", "Oct 2022"))) +
                           theme(legend.position='top') +
                           guides(color = guide_legend(nrow = 2)), 
                           p09v0 + xlab ("") + ylab(""),
                           labels = c("A", "B"),
                           nrow=1,  align = "h")

@@

ggsave(file="../4_output/figs/Fig.3 Immunity.pdf",
       plot = p09all,
        width=2400, height=2000, 
        dpi=300, units ="px")
```
## 5. Additional analysis for manuscript

####  5.1 Some desc analysis
```{r}
# median age in CBS Alberta and APL
summary(CBSs_AB$age) 
summary(APLs$age) # age=0?

@@

Among samples from our auxiliary datasets tested with both Abbott and Roche Anti-N assays, the COI was lower with Abbott than with Roche in X% of samples
```{r}
# This involves directly comparison btw Roche anti-N and Abbott-anti-N

@@

#### 5.2 PCR-verified test positivity:
We did this analysis as suggested by the reviewer-1
```{r}
# How many test results are previously PCR-verified?

@@

# Abbott Anti-S-RBD
RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  count(`RBD-IgGII_INTERP`)
```
```{r}
# Stratified analysis
RFD4682_ev2$DAYS_SINCE_NAAT<-as.numeric(RFD4682_ev2$DAYS_SINCE_NAAT)
hist(RFD4682_ev2$DAYS_SINCE_NAAT)

@@

RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  group_by(NAAT_month_cat) %>%
  count(`N-IgG_INTERP`)
```
#### 5.3 Modify the legends and y-axis labels of figure 1
```{r}
# Check the codes in qmd script 2.3
# use key words "Fig.1" "Fig.2" to search
```

>>>>

#### 2.2 Roche-S --\> Abbott-S transformation

This part is skipped and not mentioned in the manuscript.

@@

```{r}
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, group=data, colour=data)) +
  geom_point(aes(colour=data)) +
  geom_line(aes(colour=data))  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red", "purple"),
                      labels = c("AB General Pop.", "APL Vax%", 
                                 "Anti-S(+)", "Anti-S(+) N(-)")
                      )
#  scale_linetype_manual("", values=c(3,3,1,1), labels = p09.lab) 
p12
```

#### 4.2.4 Apply (S+ N-)/N- for vax% estimation

First we calculate (S+ N-)/N- among CBS samples

```{r}
table(CBSs_AB$pos_roche, CBSs_AB$pos_roche_s, useNA = 'ifany')
# good data quality, only two NAs in roche_n and roche_s

@@

Then we calculate the vax% using S(+) and (S+ & N-)/(N-)

```{r}
tbCBS_vax_est<-CBSs_AB %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            n_pos = sum(pos_roche, na.rm = T),
            s_pos = sum(pos_roche_s, na.rm = T),    
            n_vax= sum(vax_est, na.rm = T)
            )
tbCBS_vax_est$n_neg = tbCBS_vax_est$n - tbCBS_vax_est$n_pos
tbCBS_vax_est$prop.vax<-tbCBS_vax_est$n_vax/tbCBS_vax_est$n_neg

@@

#### Now redo the plot

```{r}
# Add study and adjustment
dta.vax$study[dta.vax$data %in% c("AB_general")]<-"AB_general"
dta.vax$study[dta.vax$data %in% c("CBS: S_pos_N_Neg")]<-"CBS"
dta.vax$study[dta.vax$data %in% c("APL_vax", "S_pos", "S_pos_N_Neg")]<-"APL"

@@

p12.lab <- c("AB General Pop.", "APL Vax%", 
             "APL: Anti-S(+)", "APL: Anti-S(+) N(-)", "CBS: Anti-S(+) N(-)")
p12<-dta.vax %>%
  ggplot(aes(x=samplemonth, y=prop.vax, colour=interaction(study, adj),
                                        linetype=interaction(study, adj))) +
  geom_point() +
  geom_line()  +
  theme_bw() + 
  ylim(0,1)+
  ylab("Vaccinated%") +
  scale_colour_manual("", values = c("orange", "red", "red","purple", "blue"),
                      labels = p12.lab)+
  scale_linetype_manual("", values=c(1,1,3,3,3), labels = p12.lab) +
  theme(legend.position = c(0.8,0.25))
p12
```

#### Print p12

```{r}
png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1440, res=300)
p12
dev.off()
```

#### 4.2.5 Figure S2 Panel B using transformed data

```{r}
# APL S+/N- N is transformed (Abbott-->Roche)

@@

##### Print p12 and p13 together

```{r}
p1213<-plot_grid(p12 + xlab ("Time of Sample Collection") +
                 coord_cartesian(xlim=as.yearmon(c("Oct 2020", "Feb 2023")))+
                 theme(legend.position = c(0.67,0.25)), 
                 p13 + xlab ("Time of Sample Collection") + ylab("") +
                 coord_cartesian(xlim=as.yearmon(c("Oct 2020", "Feb 2023")))+
                 theme(legend.position = c(0.65,0.2)),
                 labels = c("A", "B"),
                 nrow=1,  align = "h")
p1213

png(file="../4_output/figs/Fig.S2 Vax.png", 
    width=2400, height=1200, res=300)
p1213
dev.off()
```

##### Print p12 and p13 in pdf formats, as requested by AJE

```{r}
ggsave(file="../4_output/figs/Fig.S2 Vax.pdf", 
       width=2400, height=1200, dpi=300, units = "px", 
       plot = p1213)
```

@@

#### 4.4.1 add prop (1-dose only)

```{r}
# number of personals in each month
tb_abbsvall<-APLs %>%
  .[!(.$vax %in% c("No_ULI")),] %>%
  group_by(samplemonth) %>%
  summarise(n = n() , 
            pos_abbott_s= sum(pos_abbott_s, na.rm = T), 
            pos_abbott_s2= sum(pos_abbott_s2, na.rm = T)
            )
tb_abbsvall$n.all<-tb_abbsvall$n

@@

#### 4.4.2 Making the plot:

```{r}
p09v1a<-tb_abbsv1 %>%
  ggplot(aes(x=samplemonth)) +
  geom_point(aes(y=prop.abbs1), colour="red") +
  geom_point(aes(y=prop.abbs2), colour="red") +
  geom_line(aes(y=prop.abbs1), colour="red")  +
  geom_line(aes(y=prop.abbs2), linetype=2, colour="red")  +
  geom_line(aes(y=prop.d1), colour="darkgrey") +
  theme_bw() + 
  ylim(c(0,1)) +
  xlab("Time of Sample Collection") + 
  ylab("Proportion")+ 
  scale_linetype_manual("", values=c(1,3), labels = c("APL-Abbott (COI=50)", 
                                                      "APL-Abbott (COI=402)"))+
  theme(legend.position='top')+
  ggtitle("Stratum: 1 dose only")
p09v1a
```

#### 4.4.3 Check the time gap btw 1st dose and 2nd dose.

```{r}
APL_vax<-RFD4682_v2_vaxe %>%
  select(ULI, ULI_IDe, DOSE_COUNT, IMMUNIZATION_DATE) %>%
  .[.$DOSE_COUNT==1 | .$DOSE_COUNT==2, ]

@@

summary(as.numeric(APL_vax$time_gap))
```

## 4.5 Adding the proportion of each stratum

@@

#### The unvaccinated stratum

```{r}
# add proportion of the stratum into the data table
tb_abbsv0<-merge(tb_abbsv0, select(tb_abbsv1, samplemonth, n.all),
      by="samplemonth", all.x=T, all.y=F)
tb_abbsv0<-tb_abbsv0 %>%
  mutate(
    prop.stra = n/n.all
  )
p09v0.lab<-c("APL (manufacturer cutoff)", "APL (PRNT-50 cutoff)",
             "Proportion of the stratum")

@@

#### The 1-dose only stratum

```{r}
tb_abbsv1_melt<-reshape2::melt(select(tb_abbsv1, -c(2:4,7)), 
                        id.vars = c("samplemonth"),
                        variable.name = "COI",
                        value.name = "prop")

@@

#### The 2-dose+ stratum

```{r}
tb_abbsv2<-tb_abbsv2[-1,]

@@

### merge all four into one figue panel

```{r}
rm(p09v0a, p09v1a, p09v2a)
```

#### Figure panel

```{r}
# use cowplot package
p09.upper<-cowplot::plot_grid(p09c + ylab("Proportion") + xlab ("") +
                           coord_cartesian(xlim=as.yearmon(c("Jan 2021", "Oct 2022"))) +
                           theme(legend.position='top') +
                           guides(color = guide_legend(nrow = 2)), 
                           p09v0 + xlab ("") + ylab(""),
                           labels = c("A", "B"),
                           nrow=1,  align = "h")

@@

ggsave(file="../4_output/figs/Fig.3 Immunity.pdf",
       plot = p09all,
        width=2550, height=2000, 
        dpi=300, units ="px")
```

## 5. Additional analysis for manuscript

#### 5.1 Some desc analysis

```{r}
# median age in CBS Alberta and APL
summary(CBSs_AB$age) 
summary(APLs$age) # age=0?

@@

Among samples from our auxiliary datasets tested with both Abbott and Roche Anti-N assays, the COI was lower with Abbott than with Roche in X% of samples

```{r}
# This involves directly comparison btw Roche anti-N and Abbott-anti-N

@@

#### 5.2 PCR-verified test positivity:

We did this analysis as suggested by the reviewer-1

```{r}
# How many test results are previously PCR-verified?

@@

# Abbott Anti-S-RBD
RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  count(`RBD-IgGII_INTERP`)
```

```{r}
# Stratified analysis
RFD4682_ev2$DAYS_SINCE_NAAT<-as.numeric(RFD4682_ev2$DAYS_SINCE_NAAT)
hist(RFD4682_ev2$DAYS_SINCE_NAAT)

@@

RFD4682_ev2 %>%
  filter(NAAT_STATUS == "Positive") %>%
  group_by(NAAT_month_cat) %>%
  count(`N-IgG_INTERP`)
```

#### 5.3 Modify the legends and y-axis labels of figure 1

```{r}
# Check the codes in qmd script 2.3
# use key words "Fig.1" "Fig.2" to search
```
